The effect of linking assumptions and number of response options on inferred scalar implicature rate

### 2018-04-29 writeup.tex
added writeup files

Because the basic reconstruction of the inference is much more easily
characterized for scalar implicatures than for other implicatures,
scalar implicatures have served as a test bed for many questions in
experimental pragmatics, including, but not limited to: 1. Are scalar
inferences default inferences (in the sense of default as arising unless
blocked by marked contexts \textbf{XXX horn, levinson, degen2015)}? 2.
Are scalar inferences default inferences (in the sense that they are
computed automatically in online processing and only cancelled by
context in a second effortful step if required by context)
({\textbf{???}}; Bott \& Noveck, 2004; Breheny et al., 2006; Grodner et
al., 2010; Huang \& Snedeker, 2009)? 3. What are the (linguistic and
extra-linguistic) factors that affect whether a scalar implicature is
derived {[}({\textbf{???}});DegenTanenhaus2015; DegenTanenhaus2016;
Degen2015; DegenGoodman2014; BergenGrodner2012; Breheny2006;
FergusonBreheny2013;DeMarneffeTonhauser;DeNeys2007;Bonnefon{]}? 4. At
what age do children acquire the ability to compute implicatures
{[}Noveck2001; Reboul; Papafragou; Barner; Frank; Musolino{]}?
\textbf{XXX fill in refs}

CONTINUE HERE: motivation for examining implicature rate assumptions:

\begin{itemize}
\item
  surging interest in differences in implicature rates (eg van tiel,
  dgen tanenhaus)
\item
  implicature rates serve as basis for claims about online processing
  (bott \& noveck, degen tanenhaus)
\item
  implicature rates serve as basis for claims about children (bishop
  katsos, barner, frank)
\item
  add ({\textbf{???}})
\item
  ({\textbf{???}}) for investigations of scalar adjectives, and
  ({\textbf{???}})
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  In a truth-value judgment task, how do we know whether an
  interpretation is literal or the result of an implicature computation?
\end{itemize}

Explain the setup * the speaker produces weaker alternative from the
scale * the facts are such that the stronger alternative is true

Traditional Linking Hypotheses: * If an implicature is calculated, the
participant chooses a Non-True/Non-Right response * If an implicature is
calculated, the participant chooses the Wrong/False response * If an
implicature is calculated, the participant chooses the lower end of the
scale (2: wrong/False, 3: wrong, 4: wrong/kinda-wrong, 5:
wrong/kinda-wrong)

Questions: * Do these linking hypotheses give us different measures of
implicature computation? * If they do differ, which one is most stable?

\section{Background}\label{background}

\begin{itemize}
\item
  discussing the ways people in tha past have measured the
  \enquote{implicature rate}.
\item
  it seems like the literature takes the n(not-True)/n(Total) as the
  proporition of responses caused by implicature calculation
\item
  BUT, I remember that Jesse Snedeker said it's NOT n(not-True)/n(Total)
  but it is n(False)/n(Total)
\item
  However, this is probably not a consensus in the field because Katsos
  \& Bishop consider the mid-point response \enquote{big} on the scale
  small-big-huge (strawberry) to be the result of implicture caculation
\item
  what is the most common measure of \enquote{implicature rate} in the
  literature? Binary True/False: Noveck 2001, Chemla \& Spector 2011,
  Ternary: Katsos \& Bishop 2011
\end{itemize}

\section{Methods}\label{methods}

\subsection{Participants}\label{participants}

@@

\subsection{Procedure}\label{procedure}

The study was administered online and through Amazon Mechanical Turk.
Participants were introduced to a set of cards with pictures of one or
two animals (Figure \ref{fig:stimuli}). They were told that a
blindfolded fictional character called Bob is going to guess what
animals are on the card. In each trial, participants saw a card as well
as a sentence representing Bob's guess. For example they saw a card with
a cat on it and read the sentence \enquote{there is a cat on the card.}
The study ended after 24 trials. At the end participants were asked
about their\\
You can access and view \href{}{the online study here}.

\subsection{Design and Materials}\label{design-and-materials}

@@

\subsection{Pre-registered Analysis}\label{pre-registered-analysis}

@@

\section{Results}\label{results}

@@

\section{Analysis}\label{analysis}

@@

\section{Modeling}\label{modeling}

@@

\hypertarget{refs}{}
\hypertarget{ref-bott2004}{}
Bott, L., \& Noveck, I. (2004). Some utterances are underinformative:
The onset and time course of scalar inferences. \emph{Journal of Memory
and Language}, \emph{51}(3), 437--457.
doi:\href{https://doi.org/10.1016/j.jml.2004.05.006}{10.1016/j.jml.2004.05.006}

@@

@@

@@

@@

@@

>>>>

Because the basic reconstruction of the inference is much more easily
characterized for scalar implicatures than for other implicatures,
scalar implicatures have served as a test bed for many questions in
experimental pragmatics, including, but not limited to:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Are scalar inferences default inferences, in the sense that they arise
  unless blocked by (marked) contexts (Degen, 2015; Horn, 1984;
  Levinson, 2000)?
\item
  Are scalar inferences default inferences, in the sense that they are
  computed automatically in online processing and only cancelled by
  context in a second effortful step if required by context) {[}Bott and
  Noveck (2004);Breheny et al. (2006);Degen and Tanenhaus (2016);Grodner
  et al. (2010);Huang and Snedeker (2009);Politzer-Ahles and Fiorentino
  (2013);Tomlinson2013{]}?
\item
  What are the (linguistic and extra-linguistic) factors that affect
  whether a scalar implicature is derived {[}Zondervan (2010);Degen and
  Tanenhaus (2015); Degen and Tanenhaus (2016); Degen (2015); Degen and
  Goodman (2014); Bergen and Grodner (2012); Breheny et al. (2006);
  Breheny, Ferguson, and Katsos (2013);Marneffe and Tonhauser (2016);De
  Neys and Schaeken (2007);Bonnefon, Feeney, and Villejoubert
  (2009);Chemla2011;Potts2015{]}?
\item
  How much diversity is there across implicature types, and within
  scalar implicatures across scale types, in whether or not an
  implicature is computed (Doran, Ward, Larson, McNabb, \& Baker, 2012;
  Tiel, Miltenburg, Zevakhina, \& Geurts, 2014)?
\item
  At what age do children acquire the ability to compute implicatures
  (Barner, Brooks, \& Bale, 2011; Katsos \& Bishop, 2011; Frank;
  Musolino, 2004; Noveck, 2001; Papafragou \& Tantalou, 2004)?
\end{enumerate}

In addressing all of these questions, it has been crucial to obtain
estimates of \textbf{implicature rates}. For 1., implicature rates from
experimental tasks can be taken to inform whether scalar implicatures
should be considered default inferences. For 2., processing measures on
responses that indicate implicatures can be compared to processing
measures on responses that indicate literal interpretations. For 3.,
contextual effects can be examined by comparing implicature rates across
contexts. For 4., implicature rates can be compared across scales (or
across implicature types). For 5., implicature rates can be compared
across age groups.

A standard measure that has stood proxy for implicature rate across many
studies is the proportion of \enquote{pragmatic} judgments in
truth-value judgment paradigms {[}Bott and Noveck (2004);Noveck
(2001);Noveck and Posada (2003);Chemla and Spector (2011);Geurts and
Pouscoulous (2009);Degen and Tanenhaus (2015);De Neys and Schaeken
(2007);Degen2014{]}. In these kinds of tasks, participants are provided
a set of facts, either presented visually or via their own knowledge of
the world. They are then asked to judge whether a sentence intended to
describe those facts is true or false (or alternatively, whether it is
right or wrong, or they are asked whether they agree or disagree with
the sentence). The crucial condition for assessing implicature rates in
these kinds of studies typically consists of a case where the facts are
such that the stronger alternative is true and the target utterance is
thus also true but underinformative. For instance, Bott and Noveck
(2004) asked participants to judge sentences like \enquote{Some
elephants are mammals}, when world knowledge dictates that all elephants
are mammals. Similarly, Degen and Tanenhaus (2015) asked participants to
judge sentences like \enquote{You got some of the gumballs} in
situations where the visual evidence indicated that the participant
received all the gumballs from a gumball machine. In these kinds of
scenarios, the story goes, if a participant responds \enquote{FALSE},
that indicates that they computed a scalar implicature, eg to the effect
of \enquote{Not all elephants are mammals} or \enquote{You didn't get
all of the gumballs}, which is (globally or contextually) false. If
instead a participant responds \enquote{TRUE}, that is taken to indicate
that they interpreted the utterance literally as `Some, and possibly
all, elephants are mammals' or \enquote{You got some, and possibly all,
of the gumballs}.

Given the centrality of the theoretical notion of \enquote{implicature
rate} to much of experimental pragmatics, there is to date a surprising
lack of discussion of the basic assumption that it is adequately
captured by the proportion of FALSE responses in truth-value judgment
tasks (but see ({\textbf{???}}); Geurts and Pouscoulous (2009); Degen
and Goodman (2014); Katsos and Bishop (2011)). Indeed, the scalar
implicature acquisition literature was shaken up when Katsos and Bishop
(2011) showed that simply by introducing an additional response option,
children started looking much more pragmatic than had been previously
observed in a binary judgment paradigm. ({\textbf{???}}) allowed
children to distribute 1, 2, or 3 strawberries to a puppet depending on
\enquote{how good the puppet said it}. The result was that children gave
on average fewer strawberries to the puppet when he produced
underinformative utterances compared to when he produced literally true
and pragmatically felicitous utterances, suggesting that children do, in
fact, display pragmatic ability even at ages when they had previously
appeared not to.

But this raises an important question: in truth-value judgment task, how
do we know whether an interpretation is literal or the result of an
implicature computation? The binary choice task typically used is
appealing in part because it allows for a direct mapping from response
options -- TRUE and FALSE -- to interpretations -- literal and
pragmatic. That the seeming simplicity of this mapping is illusory
becomes apparent once a third response option is introduced, as in the
Katsos and Bishop (2011) case. How is the researcher to interpret the
intermediate option? Katsos and Bishop (2011) grouped the intermediate
option with the negative endpoint of the scale for the purpose of
categorizing judgments as literal vs.~pragmatic. But it seems just as
plausible that they could have grouped it with the positive endpoint of
the scale and taken the hard line that only truly FALSE responses
constitute a full-fledged implicature. The point here is that there has
been remarkably little consideration of \textbf{linking functions}
between behavioral measures and theoretical constructs in experimental
pragmatics, a problem in many subfields of psycholinguistics
({\textbf{???}}). We argue that it is time to engage more seriously with
these issues.

We begin by reporting an experiment that addresses the following
question: do the number of response options provided in a truth-value
judgment task and the way that responses are grouped into pragmatic
(\enquote{SI}) and literal (\enquote{no SI}) change inferences about
scalar implicature rates? Note that this way of asking the question
presupposes two things: first, that whatever participants are doing in a
truth-value judgment task, the behavioral measure can be interpreted as
providing a measure of \textbf{interpretation}. And second, that
listeners either do or do not compute an implicature on any given
occasion. In the Discussion we will discuss both of these issues. First,
following Degen and Goodman (2014), we will offer some remarks on why
truth-value judgment tasks are better thought of as measuring
participants' estimates of speakers' \textbf{production} probabilities.
This will suggest a completely different class of linking functions. And
second, we discuss an alternative conception of scalar implicature as a
probabilistic phenomeonen, a view that has recently rose to prominence
in the subfield of probabilistic pragmatics. This alternative conception
of scalar implicature, we argue, affords developing and testing
quantitative linking functions in a rigorous and motivated way.

Consider a setup in which a listener is presented a card with a
depiction of either one or two animals (see the figure below for an
example). As in a standard truth-value judgment task, the listener then
observes an underinformative utterance about this card (e.g.,
\enquote{There is a cat or a dog on the card}) and is asked to provide a
judgment on a scale from 2 to 5 response options, with endpoints
\enquote{wrong} and \enquote{right}. In the binary case, this reproduces
the standard truth-value judgment task. \textbf{XXX say briefly sth
about wrong/right vs true/false and agree/disagree}. The figure below
exemplifies (some of) the researcher's options for grouping responses.
Under what we will call the \enquote{Strong link} assumption, only the
negative endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Under the \enquote{Weak link}
assumption, in contrast, any response that does not correspond to the
positive endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Intermediate grouping schemes are also
possible, but these are the ones we will consider here. Note that for
the binary case, the Weak and Strong link return the same categorization
scheme, but for any number of response options greater than 2, the Weak
and Strong link can in principle lead to differences in inferences about
implicature rate.

@@

Let's examine an example. Assume three response options (wrong, neither,
right). Assume further that a third of participants each gave each of
the three responses, i.e., the distributions of responses is 1/3, 1/3,
and 1/3. Under the Strong link, we infer that this task yielded an
implicature rate of 2/3. Under the Weak link, we infer that this task
yielded an implicature rate of 1/3. This is quite a drastic difference
if we are for instance interested in whether scalar implicatures are
inference defaults and we would like to interpret an implicature rate of
above an arbitrary threshold (e.g., 50\%) as evidence for such a claim.
Under the Strong link, we would conclude that scalar implicatures are
not defaults. Under the Weak link, we would conclude that they are.

In the experiment reported in the following section, we presented
participants with exactly this setup. Different groups of participants
were presented with different numbers of response options. We
categorized their responses according to the Weak and the Strong link
and tested whether number of response options and categorization scheme
leads to different conclusions about implicature rates.

\section{Experiment}\label{experiment}

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

@@

\subsubsection{Procedure}\label{procedure}

The study was administered online through Amazon Mechanical Turk.
Participants were introduced to a set of cards with pictures of one or
two animals (Figure \ref{fig:stimuli}). They were told that a
blindfolded fictional character called Bob is going to guess what
animals are on the card. On each trial, participants saw a card as well
as a sentence representing Bob's guess. For example, they saw a card
with a cat on it and read the sentence \enquote{There is a cat on the
card.} The study ended after 24 trials. At the end participants
optionally provided demographic information. You can access and view
\href{**XXX\%20masoud\%20insert\%20link**}{the online study here}.

\subsubsection{Design and Materials}\label{design-and-materials}

@@

\subsection{Results}\label{results}

@@

@@

\subsubsection{Analysis}\label{analysis}

@@

@@

\hypertarget{refs}{}
\hypertarget{ref-Barner2011}{}
Barner, D., Brooks, N., \& Bale, A. (2011). Accessing the unsaid: the
role of scalar alternatives in children's pragmatic inference.
\emph{Cognition}, \emph{118}(1), 84--93.
doi:\href{https://doi.org/10.1016/j.cognition.2010.10.010}{10.1016/j.cognition.2010.10.010}

\hypertarget{ref-Bergen2012}{}
Bergen, L., \& Grodner, D. J. (2012). Speaker knowledge influences the
comprehension of pragmatic inferences. \emph{Journal of Experimental
Psychology. Learning, Memory, and Cognition}, \emph{38}(5), 1450--60.
doi:\href{https://doi.org/10.1037/a0027850}{10.1037/a0027850}

\hypertarget{ref-Bonnefon2009}{}
Bonnefon, J.-F., Feeney, A., \& Villejoubert, G. (2009). When some is
actually all: scalar inferences in face-threatening contexts.
\emph{Cognition}, \emph{112}(2), 249--58.
doi:\href{https://doi.org/10.1016/j.cognition.2009.05.005}{10.1016/j.cognition.2009.05.005}

\hypertarget{ref-Bott2004}{}
Bott, L., \& Noveck, I. (2004). Some utterances are underinformative:
The onset and time course of scalar inferences. \emph{Journal of Memory
and Language}, \emph{51}(3), 437--457.
doi:\href{https://doi.org/10.1016/j.jml.2004.05.006}{10.1016/j.jml.2004.05.006}

\hypertarget{ref-Breheny2013}{}
Breheny, R., Ferguson, H. J., \& Katsos, N. (2013). Taking the epistemic
step: Toward a model of on-line access to conversational implicatures.
\emph{Cognition}, \emph{126}(3), 423--40.
doi:\href{https://doi.org/10.1016/j.cognition.2012.11.012}{10.1016/j.cognition.2012.11.012}

@@

\hypertarget{ref-Chemla2011}{}
Chemla, E., \& Spector, B. (2011). Experimental Evidence for Embedded
Scalar Implicatures. \emph{Journal of Semantics}, \emph{28}(3),
359--400.

\hypertarget{ref-DeNeys2007}{}
De Neys, W., \& Schaeken, W. (2007). When People Are More Logical Under
Cognitive Load - Dual Task Impact on Scalar Implicature.
\emph{Experimental Psychology}, \emph{54}(2), 128--133.
doi:\href{https://doi.org/10.1027/1618-3169.54.2.128}{10.1027/1618-3169.54.2.128}

\hypertarget{ref-Degen2015}{}
Degen, J. (2015). Investigating the distribution of 'some' (but not
'all') implicatures using corpora and web-based methods. \emph{Semantics
and Pragmatics}, \emph{8}(11), 1--55.
doi:\href{https://doi.org/10.3765/sp.8.11}{10.3765/sp.8.11}

\hypertarget{ref-Degen2014}{}
Degen, J., \& Goodman, N. D. (2014). Lost your marbles? The puzzle of
dependent measures in experimental pragmatics. \emph{Proceedings of the
36th Annual Conference of the Cognitive Science Society}, 397--402.

@@

\hypertarget{ref-DegenTanenhaus2016}{}
Degen, J., \& Tanenhaus, M. K. (2016). Availability of Alternatives and
the Processing of Scalar Implicatures: A Visual World Eye-Tracking
Study. \emph{Cognitive Science}, \emph{40}(1), 172--201.
doi:\href{https://doi.org/10.1111/cogs.12227}{10.1111/cogs.12227}

\hypertarget{ref-Doran2012}{}
Doran, R., Ward, G., Larson, M., McNabb, Y., \& Baker, R. E. (2012). A
novel experimental paradigm for distinguishing between what is said and
what is implicated. \emph{Language}, \emph{88}, 124--154.

@@

\hypertarget{ref-horn1984}{}
Horn, L. (1984). Toward a new taxonomy for pragmatic inference: Q-based
and R-based implicature. In D. Schiffrin (Ed.), \emph{Meaning, form, and
use in context: Linguistic applications} (pp. 11--42). Washington:
Georgetown University Press.

@@

\hypertarget{ref-Katsos2011}{}
Katsos, N., \& Bishop, D. V. M. (2011). Pragmatic tolerance:
implications for the acquisition of informativeness and implicature.
\emph{Cognition}, \emph{120}(1), 67--81.
doi:\href{https://doi.org/10.1016/j.cognition.2011.02.015}{10.1016/j.cognition.2011.02.015}

\hypertarget{ref-levinson2000}{}
Levinson, S. C. (2000). \emph{Presumptive Meanings - The Theory of
Generalized Conversational Implicature}. MIT Press.

\hypertarget{ref-DeMarneffe2017}{}
Marneffe, M.-C. de, \& Tonhauser, J. (2016). Inferring meaning from
indirect answers to polar questions: The contribution of the
rise-fall-rise contour. In E. Onea, M. Zimmermann, \& K. von Heusinger
(Eds.), \emph{Questions in discourse}. Leiden: Brill Publishing.

\hypertarget{ref-Musolino2004}{}
Musolino, J. (2004). The semantics and acquisition of number words:
integrating linguistic and developmental perspectives. \emph{Cognition},
\emph{93}(1), 1--41.
doi:\href{https://doi.org/10.1016/j.cognition.2003.10.002}{10.1016/j.cognition.2003.10.002}

\hypertarget{ref-Noveck2001}{}
Noveck, I. (2001). When children are more logical than adults:
experimental investigations of scalar implicature. \emph{Cognition},
\emph{78}(2), 165--188. Retrieved from
\url{http://www.ncbi.nlm.nih.gov/pubmed/11074249}

@@

\hypertarget{ref-Noveck2003}{}
Noveck, I., \& Posada, A. (2003). Characterizing the Time Course of an
Implicature: an Evoked Potentials Study. \emph{Brain and Language},
\emph{85}(2), 203--210.
doi:\href{https://doi.org/10.1016/S0093-934X(03)00053-1}{10.1016/S0093-934X(03)00053-1}

\hypertarget{ref-Papafragou2004}{}
Papafragou, A., \& Tantalou, N. (2004). Children's Computation of
Implicatures. \emph{Language Acquisition}, \emph{12}(1), 71--82.

\hypertarget{ref-Politzer-Ahles2013}{}
Politzer-Ahles, S., \& Fiorentino, R. (2013). The Realization of Scalar
Inferences: Context Sensitivity without Processing Cost. \emph{PLoS
ONE}, \emph{8}(5).
doi:\href{https://doi.org/10.1371/journal.pone.0063943}{10.1371/journal.pone.0063943}

\hypertarget{ref-VanTiel2014}{}
Tiel, B. van, Miltenburg, E. van, Zevakhina, N., \& Geurts, B. (2014).
Scalar diversity. \emph{Journal of Semantics}.
doi:\href{https://doi.org/10.1093/jos/ffu017}{10.1093/jos/ffu017}

\hypertarget{ref-Zondervan2010}{}
Zondervan, A. (2010). \emph{Scalar implicatures or focus: an
experimental approach} (PhD thesis). Universiteit Utrecht, Amsterdam.
